Equivalence Testing and the Second Generation P-Value

### 2018-08-24 manuscript.Rmd
Typo

abstract: |
  To move beyond the limitations of null-hypothesis tests, statistical approaches have been developed where the observed data is compared against a range of values that are equivalent to the absence of a meaningful effect. Specifying a range of values around zero allows researchers to statistically reject the presence of effects large enough to matter, and prevents practically insignificant effects from being interepreted as a statistically significant difference. We compare the behavior of the recently proposed second generation *p*-value [@blume_second-generation_2018] with the more established Two One-Sided Tests (TOST) equivalence testing procedure [@schuirmann_comparison_1987]. We show that the two approaches yield almost identical results under optimal conditions. Under suboptimal conditions (e.g., when the confidence interval is wider than the equivalence range, or when confidence intervals are asymmetric) the second generation *p*-value becomes difficult to interpret as a descriptive statistic. The second generation *p*-value is interpretable in a dichotomous manner (i.e., when the SGPV equals 0 or 1 because the confidence intervals lies completely within or outside of the equivalence range), but this dichotomous interpretation does not require calculations. We conclude that equivalence tests yield more consistent *p*-values, distinguish between datsets that yield the same second generation *p*-value, and allow for easier control of Type I and Type II error rates.

>>>>

abstract: |
  To move beyond the limitations of null-hypothesis tests, statistical approaches have been developed where the observed data is compared against a range of values that are equivalent to the absence of a meaningful effect. Specifying a range of values around zero allows researchers to statistically reject the presence of effects large enough to matter, and prevents practically insignificant effects from being interepreted as a statistically significant difference. We compare the behavior of the recently proposed second generation *p*-value [@blume_second-generation_2018] with the more established Two One-Sided Tests (TOST) equivalence testing procedure [@schuirmann_comparison_1987]. We show that the two approaches yield almost identical results under optimal conditions. Under suboptimal conditions (e.g., when the confidence interval is wider than the equivalence range, or when confidence intervals are asymmetric) the second generation *p*-value becomes difficult to interpret as a descriptive statistic. The second generation *p*-value is interpretable in a dichotomous manner (i.e., when the SGPV equals 0 or 1 because the confidence intervals lies completely within or outside of the equivalence range), but this dichotomous interpretation does not require calculations. We conclude that equivalence tests yield more consistent *p*-values, distinguish between datasets that yield the same second generation *p*-value, and allow for easier control of Type I and Type II error rates.
